Product graphics

### 2009-12-06 product-graphics.tex
Minor tweaking

A pie chart is a popular method of displaying proportions, but is not a rectangular partition. However, it turns out that there is a simple relationship between the bar charts and pie charts: an pie chart is an hspine drawn in polar coordinates with the x coordinate mapped to angle and the y coordinate to radius. Many other displays that use circulate arcs turn out to be special cases of product plots, drawn in polar coordinates:

@@

  \item Bullseye chart, hspine.

  \item Doughnut plot, vspine + hspine.

  \item Circular bar chart or racetrack plot, vbar.

  \item Infoslices \citep{andrews:1998}, nested vbar.  But only use half of the disk, and are specialised for highly nested data.

@@

\noindent Eight variants are displayed Figure~\ref{fig:polar}.

@@

A number of non-rectangular treemaps have been proposed, such as the circular treemaps of \citet{wetzel:2008}, the space-filling curves of \citet{wattenberg:2005} and the voronoi treemaps of \citet{balzer:2005}.  However, because it is such much more difficult to compare the areas of arbitrary polygons, these approaches tend to be attractive rather than useful.

@@

% Pseudo 2d, where you take a long 1d structure and wrap it into 2d.
% 
% Pseduo 1d, where you make a new variable from multiple variables.

@@

We have assumed the probabilities represent counts, but without loss of generality we can use any set of non-negative, additive weights instead. Some of the applications of such weighted plots are described in \citet{unwin:2007}. In the happy dataset, the {\sf wtssall} variable gives analytic survey weights. Figure~\ref{fig:weighting} shows the difference between the weighted and unweighted distributions of age and sex. The distribution is subtly different.

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{wt-count}%
    \includegraphics[width=0.5\linewidth]{wt-wtssall}%
  \caption{(Left) Unweighted distribution (counts).  (Right) Weighted by probability weights.}
  \label{fig:weighting}
\end{figure}

@@

Due to the hierarchical nature of these plots, it's difficult to label them in an informative way.  We take a pragmatic approach, and label a single horizontal and single vertical split, where possible.  Understanding the composition of these plots is most easy when:

@@

\begin{itemize}
  \item colour (map of the market)
  \item texture (sieve plots)
  \item photographs
  \item text (tables)
  \item embedded plots (time series in lab escape)
\end{itemize}

>>>>

A pie chart is a popular method of displaying proportions, but is not a rectangular partition. However, it turns out that there is a simple relationship between product plots and pie charts: an pie chart is an hspine drawn in polar coordinates with the x coordinate mapped to angle and the y coordinate to radius. Many other displays that use circular arcs turn out to be special cases of product plots, drawn in polar coordinates:

@@

  \item Bullseye chart: hspine.

  \item Doughnut plot: vspine + hspine.

  \item Circular bar chart or racetrack plot: vbar.

  \item Infoslices \citep{andrews:1998}, nested vbar. But they only use half of the disk, and are specialised for highly nested data.

@@

\noindent Eight variants are displayed in Figure~\ref{fig:polar}.

@@

A number of non-rectangular treemaps have been proposed, such as the circular treemaps of \citet{wetzel:2008}, the space-filling curves of \citet{wattenberg:2005} and the voronoi treemaps of \citet{balzer:2005}.  However, because it is so much more difficult to compare the areas of arbitrary polygons, these approaches tend to be attractive rather than useful.

@@

On top of this basic framework, there are a few variations designed for dealing with special types of data: missing data, weighted data, and continuous data.  These variations are described below.

@@

We have assumed that the probabilities represent counts, but without loss of generality, we can use any set of non-negative, additive weights. For example, in the happy dataset, the {\sf wtssall} variable gives analytic survey weights. Figure~\ref{fig:weighting} shows the difference between the weighted and unweighted distributions of age and sex. The distribution is subtly different. \citep{unwin:2007} describes the applications of weighted plots in more detail.

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{wt-count}%
    \includegraphics[width=0.5\linewidth]{wt-wtssall}%
  \caption{Joint distribution of age and sex (\key{male} male, \key{female} female). (Left) counts and (right) probability weights.}
  \label{fig:weighting}
\end{figure}

@@

Due to the hierarchical nature of these plots, it's difficult to label them in an informative way. We take a pragmatic approach, and label a single horizontal and single vertical split, where possible, using colour to identify a third variable, if necessary.

Understanding the composition of these plots is most easy when:

@@

\begin{itemize}
  \item colour (map of the market)
  \item texture, such as in sieve plots \citep{friendly:2000}
  \item photographs \citep{bederson:2001}
  \item text (tables)
  \item embedded plots (time series in lab escape)
\end{itemize}
